---
title: "40-model"
output: html_notebook
---

I created some features that create groups based upon a significant threshold in the column, so now I can use this to group by and summarize other measures if I choose to.
```{r TestFunctions}
PimaIndiansDiabetes2 %>%
std_variable("glucose")

PimaIndiansDiabetes2 %>%
  obese("mass")

PimaIndiansDiabetes2 %>%
  at_risk_age("age")
```

The functions both appear to be working correctly new varaibles are construct and arranged properly in the dataset. I now want to create the augmented dataframe with the new features so that it is ready for modeling. I will also move the target variable "diabetes" to the front of the dataframe, followed by the mutated varaibles and the inital varaibles used to contruct the new features.
```{r ModelDF}
pid_augmented_df <- PimaIndiansDiabetes2 %>%
  std_variable("glucose") %>%
  obese("mass") %>%
  at_risk_age("age")%>%
  select(diabetes, everything())

write.csv(pid_augmented_df, "pid_df.csv")
```


I will run assert/verifies on the new feature varaibles to ensure that the values in these columns are what I expect. The standardized column for the input, in our case "glucose," should all be within 3 standard deviations from the mean, barring an extreme outlier. as for our "is_obese" logical variable we will just check to ensure that it only has a true or false value for each observation.
```{r AssertFunctions}
pid_df <- read_csv("pid_df.csv") %>%
  assert(within_bounds(-3,3), std_input)

pid_df <- read_csv("pid_df.csv") %>%
  assert(in_set(TRUE,FALSE), is_obese)

pid_df <- read_csv("pid_df.csv") %>%
  assert(in_set(TRUE,FALSE), age45)
  
```

